Integration of Multiple Constraints in ACG

### 2013-10-21 lenls/slides.tex
LENLS: First usable version of slides

\usepackage{xcolor}

@@

\title[G-ACGs]{Towards a Wide-Coverage Grammar}
\subtitle{Graphical Abstract Categorial Grammars}
\author{Ji\v{r}\'{i} Mar\v{s}\'{i}k}
\institute[INRIA Loria]
{
  INRIA Loria
  \and
  UFR Math\'{e}matiques et Informatiques \\
  Universit\'{e} de Lorraine
}
\date[June 2013]{June 25, 2013}

@@

\begin{frame}
  \frametitle{Towards a Wide-Coverage Grammar}

  \textbf{Objective:} Build discourse representations using ACGs: \\a
  wide-coverage ACG grammar.

  \vfill

  \begin{itemize}
  \item Having linguistic resources to build upon
    \begin{itemize}
    \item Frilex, a formalism-independent lexical database
    \item Frigram, a French interaction grammar
    \end{itemize}
    \vfill
  \item Debugging the grammar
    \begin{itemize}
    \item A toolkit for formulating ACGs as simple metagrammars, for
      examining how they work and for checking their consistency.
    \end{itemize}
    \vfill
  \item Using a formalism that allows for a simple grammar
    \begin{itemize}
    \item Express multiple independent constraints \\ $\implies$ graphical abstract categorial grammars.
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{ACGs and IGs}

  \begin{itemize}
  \item ACGs are based on types (IILL formulas).
  \item IGs are based on polarized tree descriptions.
    \vfill
  \item[$\Rightarrow$] Perrier has linked the two.

    \vspace{2 mm}
    \begin{columns}[c]
      \begin{column}{0.1\textwidth}\end{column}
      \begin{column}{0.4\textwidth}
        $$(NP \limp S) \limp N \limp N$$
        $$N \limp (NP \limp S) \limp N$$
      \end{column}
      \begin{column}{0.1\textwidth}$\to$\end{column}
      \begin{column}{0.2\textwidth}
        \includegraphics[height=0.3\textheight]{../diagrams/ptd.pdf}
      \end{column}
      \begin{column}{0.25\textwidth}\end{column}
    \end{columns}
    \vfill
  \item IG of Frigram use different constraints in tree descriptions.
    \begin{itemize}
    \item The deductive syntactic model of IGs is built on a wider range
      of linear logic primitives ($\otimes$, $\&$).
    \end{itemize}
  \end{itemize}
\end{frame}

@@

@@

  \begin{itemize}
  \item The three language families we defined above are\ldots
    \begin{itemize}
    \item {\ldots}increasingly constraining, $\forall \, \mathcal{G},
      u. \; \mathcal{I}_{\mathcal{G}}(u) \supseteq
      \mathcal{E}_{\mathcal{G}}(u) \supseteq
      \mathcal{P}_{\mathcal{G}}(u)$.
    \item {\ldots}increasingly expressive, $\mathcal{I} \subseteq
      \mathcal{E} \subseteq \mathcal{P}$.
    \end{itemize}
  \vfill
  \item Intrinsic languages are just the abstract languages of ACGs.
    $$\mathcal{I} = \mathcal{A}$$
  \vfill
  \item Extrinsic languages are simple extensions of ACGs that add
    intersections.
    $$\mathcal{E} = \mathcal{O}^{\mathcal{L}{\cap}}$$
  \vfill
  \item Pangraphical languages are faithful to the generative
    perspective and subsume the previous classes of languages.
    \begin{itemize}
    \item They solve both the example problems we saw before
      (\emph{meanings of grammatical sentences} and the \emph{diamond}).
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Summary of Work Done}

  \begin{itemize}
  \item Direct embedding of IGs into ACGs based on Perrier's theorem
    conjectured not to be practical.
  \vfill
  \item Created a toolkit to serve as basis for more explorations or
    implementations.
  \vfill
  \item Formalized graphical abstract categorial grammars.
    \begin{itemize}
    \item Helps solve multiple constraints elegantly.
    \item Facilitates discussion and inquiry into compositions of ACGs.
    \end{itemize}
  \end{itemize}
\end{frame}

@@

  \begin{itemize}
  \item Are extrinsic languages as expressive as pangraphical
    ones? $$\mathcal{P} \subseteq \mathcal{E}$$
  \item Are ACG object languages closed on
    intersection? $$\mathcal{O}^{\mathcal{L}{\cap}} \subseteq \mathcal{O}^{\mathcal{L}}$$
  \item Express the multiple polarized features of IGs in G-ACGs.
    \vfill
  \item Build and study a wide-coverage G-ACG.
    \vfill
  \item Apply G-ACGs to discourse.
  \end{itemize}
\end{frame}

\appendix
\newcounter{finalframe}
\setcounter{finalframe}{\value{framenumber}}

  Agreement:
    \begin{align*}
      A_{aime_1} &: NP\_\textcolor{darkgold}{NUM{=}SG} \limp NP\_\textcolor{darkgold}{NUM{=}SG} \limp S \\
      A_{aime_2} &: NP\_\textcolor{darkgold}{NUM{=}SG} \limp NP\_\textcolor{darkgold}{NUM{=}PL} \limp S
    \end{align*}

  Extraction:
    \begin{align*}
      E_{aime_1} &: NP\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}NO} \\
      E_{aime_2} &: NP\_\textcolor{darkgreen}{VAR{=}T} \limp NP\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT} \\
      E_{aime_3} &: NP\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgreen}{VAR{=}T} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT}
    \end{align*}

  Combined:
  \begin{small}
  \begin{align*}
      C_{aime_1} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}NO} \\
      C_{aime_2} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}T} \limp NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT} \\
      C_{aime_3} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}T} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT} \\
      C_{aime_4} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgold}{NUM{=}PL}\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}NO} \\
      C_{aime_5} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}T} \limp NP\_\textcolor{darkgold}{NUM{=}PL}\_\textcolor{darkgreen}{VAR{=}F} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT} \\
      C_{aime_6} &: NP\_\textcolor{darkgold}{NUM{=}SG}\_\textcolor{darkgreen}{VAR{=}F} \limp NP\_\textcolor{darkgold}{NUM{=}PL}\_\textcolor{darkgreen}{VAR{=}T} \limp S\_\textcolor{darkgreen}{EXT{=}ROOT}
  \end{align*}
  \end{small}
\end{frame}

\setcounter{framenumber}{\value{finalframe}}

>>>>

\usepackage[utf8]{inputenc}
\usepackage{xcolor}

@@

\title[G-ACGs]{Integration of Multiple Constraints in ACG}
\subtitle{Graphical Abstract Categorial Grammars}
%
\author{Jiří Maršík \and Maxime Amblard}
%
\institute{Université de Lorraine, Laboratoire lorrain de recherche en informatique et ses applications, UMR 7503, Vandoeuvre-lès-Nancy, 54500, France\\
INRIA, Villers-lès-Nancy, 54600, France\\
CNRS, Loria, Vandoeuvre-lès-Nancy, UMR 7503, 54500, France\\
\texttt{\{jiri.marsik, maxime.amblard\}@loria.fr}}

\date[October 2013]{October 28, 2013}

@@

\begin{frame}
  \frametitle{Introduction}

  \begin{itemize}
    \item Abstract Categorial Grammars provide interesting tools for managing
      discourse and dynamic semantics. \vfill
    \item Refining grammars to enforce numerous constraints breeds complexity.
      \vfill
    \item We formalize an extension of ACGs which lets us use the ACG
      apparatus and maintain independent constraints. \vfill
    \item Our goal is a large-scale grammar usable for discussing discourse.
  \end{itemize}
\end{frame}

@@

  Agreement:
    \begin{align*}
      A_{aime_1} &: NP_{\textcolor{darkgold}{NUM{=}SG}} \limp NP_{\textcolor{darkgold}{NUM{=}SG}} \limp S \\
      A_{aime_2} &: NP_{\textcolor{darkgold}{NUM{=}SG}} \limp NP_{\textcolor{darkgold}{NUM{=}PL}} \limp S
    \end{align*}

  Extraction:
    \begin{align*}
      E_{aime_1} &: NP_{\textcolor{darkgreen}{VAR{=}F}} \limp NP_{\textcolor{darkgreen}{VAR{=}F}} \limp S_{\textcolor{darkgreen}{EXT{=}NO}} \\
      E_{aime_2} &: NP_{\textcolor{darkgreen}{VAR{=}T}} \limp NP_{\textcolor{darkgreen}{VAR{=}F}} \limp S_{\textcolor{darkgreen}{EXT{=}ROOT}} \\
      E_{aime_3} &: NP_{\textcolor{darkgreen}{VAR{=}F}} \limp NP_{\textcolor{darkgreen}{VAR{=}T}} \limp S_{\textcolor{darkgreen}{EXT{=}ROOT}
}    \end{align*}

  Combined:
  \begin{small}
  \begin{align*}
      C_{aime_1} &: NP_{\textcolor{darkgold}{NUM{=}SG},\textcolor{darkgreen}{VAR{=}F}} \limp NP_{\textcolor{darkgold}{NUM{=}SG},\textcolor{darkgreen}{VAR{=}F}} \limp S_{\textcolor{darkgreen}{EXT{=}NO}} \\
      C_{aime_2} &: NP_{\textcolor{darkgold}{NUM{=}SG},\textcolor{darkgreen}{VAR{=}T}} \limp NP_{\textcolor{darkgold}{NUM{=}SG},\textcolor{darkgreen}{VAR{=}F}} \limp S_{\textcolor{darkgreen}{EXT{=}ROOT}} \\
      C_{aime_3} &: NP_{\textcolor{darkgold}{NUM{=}SG},\textcolor{darkgreen}{VAR{=}F}} \limp NP_{\textcolor{darkgold}{NUM{=}SG},\textcolor{darkgreen}{VAR{=}T}} \limp S_{\textcolor{darkgreen}{EXT{=}ROOT}} \\
      C_{aime_4} &: NP_{\textcolor{darkgold}{NUM{=}SG},\textcolor{darkgreen}{VAR{=}F}} \limp NP_{\textcolor{darkgold}{NUM{=}PL},\textcolor{darkgreen}{VAR{=}F}} \limp S_{\textcolor{darkgreen}{EXT{=}NO}} \\
      C_{aime_5} &: NP_{\textcolor{darkgold}{NUM{=}SG},\textcolor{darkgreen}{VAR{=}T}} \limp NP_{\textcolor{darkgold}{NUM{=}PL},\textcolor{darkgreen}{VAR{=}F}} \limp S_{\textcolor{darkgreen}{EXT{=}ROOT}} \\
      C_{aime_6} &: NP_{\textcolor{darkgold}{NUM{=}SG},\textcolor{darkgreen}{VAR{=}F}} \limp NP_{\textcolor{darkgold}{NUM{=}PL},\textcolor{darkgreen}{VAR{=}T}} \limp S_{\textcolor{darkgreen}{EXT{=}ROOT}
}  \end{align*}
  \end{small}
\end{frame}

@@

  \begin{itemize}
  \item The three language families we defined above are\ldots
    \begin{itemize}
    \item {\ldots}increasingly constraining, $\forall \, \mathcal{G},
      u. \; \mathcal{I}_{\mathcal{G}}(u) \supseteq
      \mathcal{E}_{\mathcal{G}}(u) \supseteq
      \mathcal{P}_{\mathcal{G}}(u)$.
    \item {\ldots}increasingly expressive, $\mathcal{I} \subseteq
      \mathcal{E} \subseteq \mathcal{P}$.
    \end{itemize}
  \vfill
  \item Intrinsic languages are just the abstract languages of ACGs.
    $$\mathcal{I} = \mathcal{A}$$
  \item Extrinsic languages are simple extensions of ACGs that add
    intersections.
    $$\mathcal{O} = \mathcal{A}^{\mathcal{L}}$$
    $$\mathcal{E} = \mathcal{A}^{\mathcal{L}{\cap}}$$
  \vfill
  \item Pangraphical languages are faithful to the generative
    perspective and subsume the previous classes of languages.
    \begin{itemize}
    \item They solve both the example problems we saw before
      (\emph{meanings of grammatical sentences} and the \emph{diamond}).
    \end{itemize}
  \end{itemize}
\end{frame}

\section{Illustration}

\begin{frame}
  \frametitle{Basic Setup}
  \begin{center}
    \includegraphics[height=0.8\textheight]{../diagrams/illustration0.pdf}
  \end{center}
\end{frame}

\begin{frame}
  \frametitle{Single Constraint}
  \begin{center}
    \includegraphics[height=0.8\textheight]{../diagrams/illustration1.pdf}
  \end{center}
\end{frame}

\begin{frame}
  \frametitle{Multiple Constraints}
  \begin{center}
    \includegraphics[height=0.8\textheight]{../diagrams/illustration2.pdf}
  \end{center}
\end{frame}

\begin{frame}
  \frametitle{Exposing Another Signature}
  \begin{center}
    \includegraphics[height=0.8\textheight]{../diagrams/illustration3.pdf}
  \end{center}
\end{frame}

\begin{frame}
  \frametitle{Constraints on Multiple Signatures}
  \begin{center}
    \includegraphics[height=0.8\textheight]{../diagrams/illustration4.pdf}
  \end{center}
\end{frame}

\begin{frame}
  \frametitle{Reflection}
  \begin{columns}[c]
    \begin{column}{0.5\textwidth}
      \begin{center}
        \includegraphics[height=0.8\textheight]{../diagrams/illustration5.pdf}
      \end{center}
    \end{column}
    \begin{column}{0.5\textwidth}
      \begin{itemize}
        \item 9 signatures
        \item 8 lexicons
        \item Constraints decomplected from each other
        \item Constraints on different signatures
        \item High level of redundancy
      \end{itemize}
    \end{column}
  \end{columns}
\end{frame}

@@

\begin{frame}
  \frametitle{Recapitulation}

  \begin{itemize}
  \item We reinterpreted ACGs by treating the diagrams as primary constructs
    instead of as artifacts.
  \vfill
  \pause
  \item This lets us handle non-arborescent diagrams in a principled manner.
  \vfill
  \pause
  \item Non-arborescence enables intersection and therefore conjunction of
    independent constraints.
  \vfill
  \pause
  \item We presented the kind of grammar we envisioned, built by collecting
    fragments described in existing research.
  \end{itemize}
\end{frame}

@@

  \begin{itemize}
  \item Are extrinsic languages as expressive as pangraphical
    ones? $$\mathcal{P} \subseteq \mathcal{E}$$
  \item Are ACG object languages closed on
    intersection? $$\mathcal{O}^{\mathcal{L}{\cap}} \subseteq \mathcal{O}^{\mathcal{L}}$$
  \item Is there a metagrammatical approach that makes refining the types of a
    signature more concise?
  \vfill
  \item Is this enough to build a satisfactory grammar of a real language?
  \end{itemize}
\end{frame}
